---
title: "Exercise 5"
output: html_notebook
---

1. Prepare the dataset: Clean the NA-observations, check the datatypes and so on. At first, let's read data.
```{r}
loans = read.csv("loans.csv", na.strings ="NA")
str(loans)
```
Let's change all data types to numeric and omit NA values.
```{r}
loans$credit.policy=as.numeric(loans$credit.policy)
loans$purpose=as.numeric(loans$purpose)
loans$fico=as.numeric(loans$fico)
loans$revol.bal=as.numeric(loans$revol.bal)
loans$inq.last.6mths=as.numeric(loans$inq.last.6mths)
loans$delinq.2yrs=as.numeric(loans$delinq.2yrs)
loans$pub.rec=as.numeric(loans$pub.rec)
loans$not.fully.paid=as.numeric(loans$not.fully.paid)

loans=na.omit(loans)
str(loans)
```
All data is now numeric and ready for the analysis.

2. Split the data into train set (75%) and test set (25%). You need a package “caTools” and its function sample.split, which creates a binary vector telling into which set each observation belongs.
```{r}
install.packages("caTools")
library(caTools)
split=sample.split(loans$not.fully.paid, SplitRatio=0.75)
treeni=subset(loans,split==TRUE)
testi=subset(loans,split==FALSE)
```
4. Form a logistic regression model with 3-4 independent variables. Try a couple and choose the best one.

Let's look at correlations of different variables in the dataset.
```{r}
cor(treeni)
```
I want to select the model so that I get good correlations with each chosen independent variable and dependent variable not.fully.paid. The independent variables that correlate the most with not.fully.paid are credit.policy (-0.15), int.rate (0.16), fico (-0.15) and inq.last.6mths(0.13). Credit.policy correlates moderately with fico (0.35), int.rate (-0.31) and quite a lot with inq.last.6mths(-0.53). Fico correlates quite a lot with with int.rate (-0.71) and moderately with inq.last.6mths (-0.18). Int.rate correlates moderately with inq.last.6mths (0.20). 

Taking into account higher correlation of credit.policy and inq.last.6mths, and that of fico and int.rate, highly correlating variables should not be included in the same model.

Let's find the third variable for the models. Revol.util correlates a little bit with not.fully.paid (0.08).It correlates with credit.policy(-0.11), fico (-0.54) and int.rate(0.46).It can be applied if fico is not present. 

Is there any other variable that correlates with not.fully.paid but does not correlate with fico less than -/+0.5? Let's take revol.bal to the model since it is not too much correlating with credit.policy (-0.19) or fico (-0.01) but somewhat with not.fully.paid(0.06). 

The three models are following:
- model1 includes credit.policy, fico, revol.bal
- model2 includes inq.last.6mths, int.rate, revol.util
- model3 includes credit.policy, int.rate, revol.util

Let's create model1.
```{r}
model1=glm(not.fully.paid ~ credit.policy+fico+revol.bal,data=treeni,family=binomial)
summary(model1)
```
Let's create model2.
```{r}
model2=glm(not.fully.paid ~ inq.last.6mths+int.rate+revol.util,data=treeni,family=binomial)
summary(model2)
```
Let's create model3.
```{r}
model3=glm(not.fully.paid ~ credit.policy+int.rate+revol.util,data=treeni,family=binomial)
summary(model3)
```
The AIC:s of the models are following:
- model1 6044.2
- model2 6035.5
- model3 6027.8

The model3 has the smallest AIC so it is selected for further analysis.

5. Test the model on the test set. Compare accuracies with different tresholds. Report.

Let's test model3 with test set.
```{r}
testPredict=predict(model3,newdata=testi, type="response")
```
Let's calculate baseline accuracy. At first, make a table.
```{r}
table(testi$not.fully.paid)
```
Baseline accuracy can now be calculated.

```{r}
acc_baseline=1999/(1999+380)
acc_baseline
```
Calculate accuracy with threshold 0.35. Create table.
```{r}
table(testi$not.fully.paid, testPredict>=0.35)
```
Calculate accuracy.
```{r}
acc_thirtyfive=(1959+22)/(1959+40+358+22)
acc_thirtyfive
```
Accuracy with threshold 0.35 is less than accuracy for the baseline model. Let's try with accuracy 0.55. At first, let's make a table.
```{r}
table(testi$not.fully.paid, testPredict>=0.55)
```
Calculate accuracy.
```{r}
acc_fiftyfive=1999/(1999+380)
acc_fiftyfive
```
With threshold 0.55, the accuracy is the same as the baseline model. Thus model3 can't beat baseline model according to these experimentations with different thresholds.
